\chapter{Statistical physics}
\minitoc

\section{Thermodynamics}

\begin{definition}[system state]
System state refers to system's thermodynamic equilibrium state, which nothing happens.
\end{definition}

\begin{remark}[steady state is not a state]
If the system is at steady state, for example, during heat, mass, and energy transport, then such state is not considered as thermodynamic equilibrium state.
\end{remark}


\begin{theorem}[first law of thermodynamics]
	$$\Delta U = W + Q$$
	where $\Delta U = U_B - U_A$ is the net change in the system's internal energy, $W$ is the work performed on the system(displacements $dX$ against force $F$), $Q$ is the heat absorbed by the system(spontaneous flow of energy via thermal contact).
\end{theorem}


\begin{remark}[heat vs. work]\cite[159]{smith2005introduction} \hfill
	\begin{itemize}
		\item Work is readily transformed into other forms of energy, the efficiency can be 100\% by elimination of friction.
		\item Heat is a form of energy intrinsically less useful and hence less valuable than an equal quantity of work or mechanical or electric energy. Conversion of heat into work or into mechanical energy never reaches 100\%. 
	\end{itemize}
\end{remark}

\begin{theorem}[second law of thermodynamics]
	$$\Delta S \geq \int_A^B \frac{dQ}{T}$$
	where $\Delta S = S_B - S_A$ is the net change in the system's entropy, $T$ is the temperature of thermal surroundings, $dQ$ is the heat absorbed by the system(spontaneous flow of energy via thermal contact).
\end{theorem}

\begin{theorem}[second law of thermodynamics, statement]\cite[159]{smith2005introduction}
No apparatus can operate in such a way that its only effect(in system and surroundings) is to convert heat absorbed by a system completely into work done by the system. 	
\end{theorem}



\begin{lemma}[isothermal processes]
	For an isothermal process where $T$ is constant, we have
	$$\Delta S \geq \frac{Q}{T} = \frac{\Delta U - W}{T}$$
	that is
	$$W \geq \Delta F = \Delta U - T\Delta S.$$
\end{lemma}




\section{Notes on Bibliography}




\printbibliography

